---
title: "Data cleaning"
date: "`r format(Sys.time(), '%F')`"
---

```{r load-libraries, message=FALSE}
library(tidyverse)
library(readxl)
library(countrycode)
```

## Collect data

### 3P Trafficking Index

[Cho et al.'s 3P Trafficking Index](http://www.economics-human-trafficking.org/data-and-reports.html) uses World Bank codes which don't always perfectly correspond with ISO3 or COW codes. We ignore some countries and make manual changes for others. 

We fix these:

- BHU (Bhutan): 760 / BTN
- KSV (Kosovo): 347 / XKK
- RMI (Marshall Islands): 983 / MHL
- HKG (Hong Kong): 715
- SRB (Serbia): 340

We ignore these:

- BES (Caribbean Netherlands): no cowcode / no ISO
- ABW (Aruba): no cowcode
- ANT (Netherlands Antilles): no cowcode
- CUW (Curacao): no cowcode
- MAC (Macao): no cowcode

```{r load-3p, message=FALSE}
cho.raw <- read_excel("data/Cho 3Ps/3P_Index_2000-2015.xlsx", skip = 2)

countries.to.ignore <- c("ABW", "ANT", "BES", "CUW", "MAC")
wb2iso <- c(BES = NA, BHU = "BTN", KSV = "XKK", RMI = "MHL")
wb2cow <- c(ABW = NA, ANT = NA, BES = NA, CUW = NA, MAC = NA,
            BHU = 760L, HKG = 715L, KSV = 347L, RMI = 983L, SRB = 340L)

cho.clean <- cho.raw %>%
  mutate(iso3 = countrycode(Code, "wb", "iso3c", custom_match = wb2iso),
         cowcode = countrycode(Code, "wb", "cown", custom_match = wb2cow)) %>%
  filter(!(Code %in% countries.to.ignore)) %>%
  select(year = Year, iso3, cowcode, 
         prosecution = Prosecution, protection = Protection,
         prevention = Prevention, p = `Overall 3P`)

cho.panel <- cho.clean %>%
  expand(cowcode, year) %>%
  left_join(cho.clean, by = c("cowcode", "year"))
```



## Combine data
