---
title: "Effect of criminalization"
date: "`r format(Sys.time(), '%F')`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.retina = 2,
                      tidy.opts=list(width.cutoff = 100),  # For code
                      options(width = 100))  # For output
```

```{r load-libraries, message=FALSE}
library(tidyverse)
library(haven)
library(stargazer)
library(broom)
library(purrr)
library(forcats)
library(viridisLite)
library(ggstance)

# By default, R uses polynomial contrasts for ordered factors in linear models
# options("contrasts") 
# So make ordered factors use treatment contrasts instead
options(contrasts=rep("contr.treatment", 2))
```

```{r load-data}
df.crim <- readRDS("data/df_crim.rds") %>%
  filter(year > 2002, year < 2016) %>%
  mutate(cowcode.i = as.factor(cowcode), year.i = as.factor(year))
```

**Our main questions**: When external actors push for legal reforms, is it just a cheap response? or is there actual change? Does it matter whether domestic pressure changes domestic laws, or is it just paper compliance?

On p. 161 in *Scorecard Diplomacy* in "Is Criminalization Meaningful or Just Cheap Talk?", we look just at averages. We can get better details than just averages.


## TODO

Variables/data to get:

- Presence/absence of action plan
- Richard Frank's data (<https://humantraffickingindicators.org/>)
- NGO density
- Years since criminalization to get cumulative effect

Modeling things:

- Model time since criminalization
- Run as logit because there's not a lot of partials
- Can we think of another example where there has been external pressure (perhaps from a treaty) to implement reforms, and where we can consider implementing legislation.
- Is there a way to do mediation analysis? Is criminalization mediating the effects of the palermo protocol or of tier pressure — or a combination?

Other issues

- Partial vs. full criminalization (OLS gets that; logit doesn't)
- Figure 6.8 from the book - look at two of the three Ps (not prosecution) as DVs in Cho’s models
- Also code if countries have an action plan or not - that might matter too - control for US effort (not the TIP score, but other measures of aid, like Wikileaks, USAID, number of trafficking NGOs in the country, etc.)
- Add case studies?
- Other ways to add additional evidence (like mediation analysis)
- Split by how bad criminalization is - marginal changes harder for richer, better countries - split by incidence? (but those are constant numbers)

It'd be really nice to show that criminalization is a mediating factor between original pressure and outcome


## Run models

For now, we just add the degree of criminalization as a continuous variable (0–2) to the original Cho OLS models.

```{r crim-effect}
model1 <- lm(prot_prev ~ prot_prev_lag + crim.level_lag + 
               corruption + polity + female.leg.prop + wecon + 
               palermo.ratified + gdp.capita_log + us.aid.pct.gdp + 
               year.i + cowcode.i, 
             data = df.crim)

model2 <- lm(prevention ~ prevention_lag + crim.level_lag + 
               corruption + polity + female.leg.prop + wecon + 
               palermo.ratified + gdp.capita_log + us.aid.pct.gdp + 
               year.i + cowcode.i, 
             data = df.crim)

model3 <- lm(protection ~ protection_lag + crim.level_lag + 
               corruption + polity + female.leg.prop + wecon + 
               palermo.ratified + gdp.capita_log + us.aid.pct.gdp + 
               year.i + cowcode.i, 
             data = df.crim)

model4 <- lm(prot_prev ~ prot_prev_lag2 + crim.level_lag2 + 
               corruption + polity + female.leg.prop + wecon + 
               palermo.ratified + gdp.capita_log + us.aid.pct.gdp + 
               year.i + cowcode.i, 
             data = df.crim)

model5 <- lm(prevention ~ prevention_lag2 + crim.level_lag2 + 
               corruption + polity + female.leg.prop + wecon + 
               palermo.ratified + gdp.capita_log + us.aid.pct.gdp + 
               year.i + cowcode.i, 
             data = df.crim)

model6 <- lm(protection ~ protection_lag2 + crim.level_lag2 + 
               corruption + polity + female.leg.prop + wecon + 
               palermo.ratified + gdp.capita_log + us.aid.pct.gdp + 
               year.i + cowcode.i, 
             data = df.crim)
```

Preliminarily it looks like criminalization matters substantially. Moving from 0 to 1 (or 1 to 2) in the preceding year is associated with a 0.14 point increase in the protection + prevention indexes and a 0.1 point increase in the protection index. The effect is positive for prevention, but not significant.

```{r display-crim-effect, results="asis"}
models <- list(model1, model2, model3, model4, model5, model6)

num.models <- length(models)

var.labs <- c("Protection + prevention (t − 1)", "Prevention (t − 1)", 
              "Protection (t − 1)", "**Criminalization (t - 1)**",
              "Protection + prevention (t − 2)", "Prevention (t − 2)", 
              "Protection (t − 2)", "**Criminalization (t - 2)**",
              "Control of corruption", "Democracy", 
              "Women legislators (%)", "Women economic rights",
              "Palermo Protocol ratified", "GDP per capita (log)", 
              "US aid (% of GDP)")

dep.var.labels <- c("Protection + Prevention", "Prevention", "Protection",
                    "Protection + Prevention", "Prevention", "Protection")

col.labs <- sprintf("(%s)", 1:num.models)

extra.lines <- list(c("Model type",
                      rep("OLS", num.models)),
                    c("Country fixed effects",
                      rep("Yes", num.models)),
                    c("Year fixed effects",
                      rep("Yes", num.models)))

title <- "Checking for the effect of criminalization"

stargazer(models, 
          type = "html", omit = c("cow", "Constant", "year"), 
          add.lines = extra.lines, title = title, covariate.labels = var.labs,
          dep.var.labels = dep.var.labels, dep.var.caption = "",
          keep.stat = c("n", "adj.rsq"),
          model.numbers = FALSE, column.labels = col.labs, 
          model.names = FALSE, dep.var.labels.include = TRUE)
```


## Estimating direct effects

### Example using Fearon and Laitin 2003

```{r}
# Panel-based lags (but tidyverse is way easier than this...)
pan.lag <- function(x, ind, lag = 1) {
  unlist(tapply(x,ind, function(x) c(rep(NA, times = lag),x[-((length(x) - lag +1):length(x))])))
}

# Load and clean data
fearon <- read_dta("data/repdata.dta")
fearon_subset <- filter(fearon, onset < 4) %>%
  mutate(instab2 = 1 * (abs(polity2l - pan.lag(polity2, ccode, 2)) > 3))

# Baseline model
baseline <- lm(onset ~  lmtnest + ncontig + Oil + ethfrac + relfrac,
               data = fearon_subset)
# summary(baseline)

# Baseline model + 
ptbias <- lm(onset ~  lmtnest + ncontig  + Oil + 
               ethfrac + relfrac + instab2, 
             data = fearon_subset)
# summary(ptbias)

# First stage (to get effect of instability)
first <- lm(onset ~ warl + gdpenl + lpop + lmtnest +
              ncontig + Oil + nwstate + instab + polity2l + ethfrac + relfrac,
            data = fearon_subset)
# summary(first)

# Second stage (CDE of ethfrac without instability)
direct <- lm(I(onset - coef(first)["instab"] * instab) ~ lmtnest + ncontig + 
               Oil + ethfrac + relfrac,
             data = fearon_subset)
# summary(direct)

coefs <- tribble(
  ~model_name, ~model,
  "Baseline", baseline,
  "With instability", ptbias,
  "Fearon and Laitin", first,
  "ACDE", direct) %>%
  mutate(tidified = model %>% map(~ tidy(., conf.int = TRUE))) %>%
  unnest(tidified)

to_plot <- coefs %>%
  filter(term == "ethfrac") %>%
  mutate(model_name = fct_rev(fct_inorder(model_name)))

ggplot(to_plot, aes(x = estimate, y = model_name, color = model_name)) +
  geom_pointrangeh(aes(xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0) +
  scale_color_viridis_d(option = "magma", end = 0.9, guide = FALSE) +
  labs(x = "Estimate", y = NULL) +
  theme_light()
```

### Our own average controlled direct effect estimation

If we want to do the same ACDE estimation, we have to distinguish between these types of variables:

- Pretreatment confounders
- Intermediate confounders
- Posttreatment variables
